---
title: "Assignment 4 Textbook Exercises"
author: "Doug Goon"
date: "April 27, 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
```


### 10.5
1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).
```{r}
is.tibble(mtcars)
```

2. Compare and contrast the following operations on a data.frame and equivalent tibble.
What is different?
Tibble doesn't allow for partial matching when subsetting. 
```{r}
df <- data.frame(abc = 1, xyz = "a")
df_tibble <- tibble::tibble(abc = 1, xyz = "a")
```

```{r}
#partial matching
df$x 
```

```{r}
#doesn't do partial matching
df_tibble$x 
```

```{r}
#becomes factor
df[, "xyz"]
```

```{r}
#returns tibble
df_tibble [,"xyz"] 
```

```{r}
#returns as factor type
df[, c("abc", "xyz")] 
```

```{r}
#return as character type
df_tibble[, c("abc", "xyz")] 
```

Why might the default data frame behaviours cause you frustration?
It might cause frustration because the functions might not work. 

3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?
```{r}
var <- "abc"
df_tibble[[var]]
```

4. Practice referring to non-syntactic names in the following data frame by
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

Extracting the variable called 1.
```{r}
annoying$`1`
```

Plotting a scatterplot of 1 vs 2.
```{r}
ggplot(data = annoying)+
  geom_point(mapping = aes(x = `1`, y = `2`))
```

Creating a new column called 3 which is 2 divided by 1.
```{r}
annoying <-mutate(annoying,
       `3` = `2`/`1`
       )
```

Renaming the columns to one, two and three.
```{r}
dplyr::rename(annoying, One =`1`, Two = `2`, Three = `3`)
```

5.
What does tibble::enframe() do? 

tibble::enframe() converts named atomic vectors or lists to two-column data frames

When might you use it?
```{r}
alphabet <- letters[1:10]
enframe(alphabet)
```

6. What option controls how many additional column names are printed at the footer of a tibble?
```{r}
?print.tbl_df
print(as_tibble(mtcars), n = 3)
```

12.6.1
```{r}
tidyr::who

who %>%
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = stringr::str_replace(code, "newrel", "new_rel")) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1)
```

3. I claimed that iso2 and iso3 were redundant with country. Confirm this claim.
```{r}
who %>% 
  select(1:3) %>% 
  sapply(function(x){length(unique(x))})
```

```{r}
who %>% select(1:3) %>%
  unite(combined, 1:3) %>%
  select(combined) %>%
  distinct() %>%
  nrow()
```
Claim confirmed.

4. For each country, year, and sex compute the total number of cases of TB. Make an informative visualisation of the data.
```{r}
who %>% 
  gather(code, value, new_sp_m014:newrel_f65, na.rm = TRUE) %>% 
  mutate(code = stringr::str_replace(code, "newrel", "new_rel")) %>%
  separate(code, c("new", "var", "sexage")) %>% 
  select(-new, -iso2, -iso3) %>% 
  separate(sexage, c("sex", "age"), sep = 1) %>%
  group_by(country, year, sex) %>%
  summarize(total_case = sum(value)) %>%
  unite(country_sex, country, sex, remove = FALSE) %>%
  ggplot() +
  geom_line(mapping = aes(x = year, y = total_case, color = sex,
                          group = country_sex))
```


